Ordinal Models

### 2024-12-07 docs/misc/discrep.qmd
update results section. And some formating improvements.

---
title: "Discrep_Patterns"
author:
- name: Thomas E. Gorman
  url: https://tegorman13.github.io/
  affiliations: 
  - name:  Communication and Cognition Lab, Purdue University, USA
    affiliation-url: https://web.ics.purdue.edu/~treimer/
lightbox: true
toc: false
code-fold: show
code-tools: true
execute:
  warning: false
---

@@

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,ggdist,gghalves,patchwork,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("Scripts/", .x, ".R"))))
theme_set(theme_nice())

>>>>

---
title: "Discrep_Patterns"
toc: false
execute:
  warning: false
---

@@

```{r}
library(pacman)
pacman::p_load(dplyr,purrr,tidyr,here,tibble, haven,
               ggplot2,gt,knitr,kableExtra,ggh4x,ggdist,gghalves,patchwork,lme4,flextable)
options(digits=2, scipen=999, dplyr.summarise.inform=FALSE)
walk(c("fun_plot"), ~ source(here::here(paste0("scripts/", .x, ".R"))))
theme_set(theme_nice())

### 2024-12-07 docs/misc/ordinal.qmd
update results section. And some formating improvements.

>>>>

marginal_effects(ordinal_model_s2_probit)

## S2 logit version
```{r}

ordinal_model_s2_logit <- brm(
  accuracy_level ~ refClass +rounded+pct_goal + (1|id)+ (1|state),
  data = s2_agg1,
  family = cumulative("logit"),
  cores = 4,
  iter = 2000,
  control = list(adapt_delta = 0.99), # Recommended for ordinal models
  prior = c(prior(normal(0, 2), class = "Intercept"),  # Priors for thresholds
            prior(normal(0, 2), class = "b")), # Priors for predictors
  file = paste0(here::here("data/model_cache",'s2_logit_add.rds')) # Cache for efficiency
)
summary(ordinal_model_s2_logit)
# Regression Coefficients:
#                    Estimate Est.Error l-95% CI u-95% CI Rhat Bulk_ESS Tail_ESS
# Intercept[1]          -2.13      0.64    -3.39    -0.86 1.00      639     1328
# Intercept[2]          -0.62      0.63    -1.89     0.63 1.00      658     1366
# Intercept[3]           3.15      0.65     1.88     4.42 1.00      717     1638
# refClassPercentage     0.83      0.75    -0.64     2.33 1.00      452     1253
# refClassUSD            1.87      0.79     0.31     3.35 1.00      508     1260
# roundedRounded        -0.66      0.18    -1.01    -0.31 1.00     4276     3095
# pct_goal15%           -0.44      0.18    -0.79    -0.10 1.00     6263     3109

pred_summary_s2 <- ordinal_model_s2_logit %>%
  epred_draws(newdata = s2_agg1, re_formula = NA) %>%
  group_by(refClass, rounded, pct_goal, .category) %>%
  summarise(
    mean_prob = mean(.epred),
    lower_ci = quantile(.epred, 0.025),
    upper_ci = quantile(.epred, 0.975),
    .groups = "drop"
  )

### logit vs. probit

loo_logit <- loo(ordinal_model_s2_logit)
loo_probit <- loo(ordinal_model_s2_probit)
loo_compare(loo_logit, loo_probit)

waic(ordinal_model_s2_logit)
waic(ordinal_model_s2_probit)

pp_check(ordinal_model_s2_logit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class - logit") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))

pp_check(ordinal_model_s2_probit, type = "bars_grouped", group="refClass", fatten = 2) +
  scale_x_continuous("Response Category", breaks = 1:4, 
            labels = c("Exact", "0.01-2%", "2.01-15%", ">15%")) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  ggtitle("Posterior Predictive Check by Reference Class - probit") +
  theme_minimal() +
  theme(
    legend.background = element_blank(),
    legend.position = "bottom",
    panel.grid.minor = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1))
